---
title: "01_explor"
author: "Puvvula"
date: "2023-01-31"
output: pdf_document
---
NOTES
2019-2020 cycle data combined with 2017-2018 cycle to build nationally representative data [pre-pandemic data] 
2021-2022 cycle not available

2019 onwards need manual access to files
examination data (file: P_OHXDEN.xpt | variable: OHX##TC); 
blood - total mercury: (file: P_PBCD.xpt | variable: LBXTHG )
blood - inorganic, ethyl, methyl mercury: (file: P_IHGEM.xpt | variables: LBXIHG, LBXBGE, LBXBGM)
urine - inorganic mercury: (file: P_UHG.xpt | variable: URXUHG)


ex: https://ehsanx.github.io/SPPH504007SurveyData/docs/importing-nhanes-to-r.html#nhanes-dataset
nhanesA coverate 1999-2018

```{r}
library(pacman)
p_load(tidyverse, haven, survey)
```

#filter women of reproductive age from NHANES demographics dataset
```{r}
nhanesDemo <- read_xpt(url(
  "https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DEMO_I.XPT")) |>
  select(SEQN, INDFMPIR, RIDAGEYR,  RIAGENDR, WTINT2YR, SDMVPSU, SDMVSTRA)

nhanesDesign <- svydesign(id      = ~SDMVPSU,
                          strata  = ~SDMVSTRA,
                          weights = ~WTINT2YR,
                          nest    = TRUE,
                          data    = nhanesDemo)
```

#downloaded NHANES data 2015-2016 cycle from : https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=Laboratory&CycleBeginYear=2015
```{r}
folder_location <- "~/Documents/nhanes/nhanes_2015_2016/"
files <- list.files(folder_location, pattern = "\\.xpt$", ignore.case = TRUE, full.names = TRUE)

for (file in files) {
  df_name <- tools::file_path_sans_ext(basename(file))
  assign(df_name, read_xpt(file))
}
```

#select only urine creatinine
```{r}
u_cr<- ALB_CR_I |>
  select(SEQN, URXUCR, URDUCRLC)

```

#creatinine adjustment for urinary biomarker concentarions
```{r}

```



#join phthlates to nhanes demographic data
```{r}
dt_join <- left_join(nhanesDemo, PHTHTE_I, 
                     by="SEQN")
```

#detection frequencies
```{r}
svytable(~dt_join$URDCNPLC, design = nhanesDesign)  |>
  as_tibble() |>
  mutate(prop= n/sum(n))
```

#quantile summary
```{r}
svyquantile(x= ~dt_join$URXMOH,
            design = nhanesDesign,
            na.rm=T,
            quantiles = c(0.25, 0.50, 0.75)) 
```



